---
title: "My Plot_ly plots"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---
```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(viridis)
library(p8105.datasets)
library(plotly)
data("instacart")
```
```

Column {data-width=650}
-----------------------------------------------------------------------

### Chart A

```{r}
veggies <-
  instacart %>% 
  filter(department == "produce") %>% 
  count(product_name, sort = TRUE) %>% 
  top_n(15) %>% 
  select(product_name)


inner_join(instacart, veggies) %>% 
  group_by(product_name, order_dow) %>% 
  count(sort = TRUE) %>% 
  plot_ly(y = ~n, x = ~order_dow, color = ~product_name, type = "bar",
          colors = "Set2") %>% 
  layout(barmode = "stack", 
         title = "15 most popular produce items",
         xaxis = list(title = "Day of Week"),
         yaxis = list(title = "Items sold"))
```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B

```{r}
instacart %>% 
  group_by(order_hour_of_day) %>% 
  filter(product_name %in% c("Limes", "Large Lemon", "Organc Yellow Onion", "Organic White Onion", "Organic Hass Avocado", "Organic Avocado")) %>% 
  count(product_name, sort = TRUE) %>% 
  plot_ly(x = ~order_hour_of_day, y = ~n, color = ~product_name, type = "scatter", mode = "line") %>% 
  layout(barmode = "stack", 
         title = "At what times are Guac Ingredients Bought?",
         xaxis = list(title = "Hour of Day"),
         yaxis = list(title = "Items sold"))
```

### Chart C

```{r}
instacart %>% 
  filter(department %in% c("produce", "babies")) %>% 
  group_by(user_id, department) %>%
  count() %>% 
  spread(key = department, value = n) %>% 
  filter(!is.na(babies) & !is.na(produce)) %>% 
  ungroup() %>% 
  sample_n(1000) %>% 
  plot_ly(x = ~produce, y = ~babies, type = "scatter", mode = "markers") %>% 
  layout(title = "Is there a relationship between baby products and produce items?",
         xaxis = list(title = "Produce Items Bought"),
         yaxis = list(title = "Baby Items Bought"))
```


